
---
output: # rmarkdown::github_document
  html_document:
  pdf_document: default
  word_document: default
title: "Assignment 3.  Working with Data Frame. Base R Style"
---

***Note*:** *This assignment practices working with Data Frame using Base R.*

***How to do it?***: 

- Open the Rmarkdown file of this assignment ([link](assignment3.Rmd)) in Rstudio. 

- Right under each question, insert  a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question. 

- `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.  

-  Publish the html file to your Githiub Page. 

***Submission***: Submit the link on Github of the assignment to Canvas under Assignment 3. 

-------

Problems
-------
 \

1. Create the following data frame

| Rank 	| Age 	| Name  	|
|:------	|:-----	|:-------	|
| 0    	| 28  	| Tom   	|
| 1    	| 34  	| Jack  	|
| 2    	| 29  	| Steve 	|
| 3    	| 42  	| Ricky 	|
```{r}
df = data.frame(Rank = c(0,1,2,3), Age = c(28,34,29,42), Name = c('Tom', 'Jack', 'Steve', 'Ricky'))
```

2. Use `read.csv` to import the Covid19 Vaccination data from WHO: [link](https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-states.csv).
```{r}
df = read.csv('https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-states.csv')
```

2. Show the names of the variables in the data

```{r}

names(df)
```


3. How many columns and rows the data have? 
```{r}

str(df)

    
```

4. How many missing values are there?  Show the missing values by columns.  What variable has the most number of missing values?
```{r}
colSums(is.na(df))

```

5. What is the class of the `date` column. Change the `date` columns to `date` type using the `as.Date` function. Show the new class of the `date` column. 
```{r}
class(df$date)
```
```{r}
df$date = as.Date(df$date)
class(df$date)

```

6. Capitalize the names of all the variables
```{r}
names(df) = c("Date", "State", "Fips", "Cases", "Deaths")
```

7. Find the average number of cases per day. Find the maximum cases a day.  
```{r}
mean(df$Cases, na.rm=TRUE)

```
```{r}
max(df$Cases, na.rm=TRUE)
```

8. How many states are there in the data?  
```{r}
length(table(df$State))

```

9. Create a new variable `weekdays` to store the weekday for each rows. 
```{r}
df$weekdays <- weekdays(df$Date)
```

10. Create the categorical variable `death2` variable taking the values as follows

- `has_death` if there is a death that day
-  `no_death` if there is no death that day
```{r}
df$death2 <- ifelse(df$Deaths==0,'has_death','no_death')

```

Find the frequency and relative frequency of `no_death` and `has_death`. 
```{r}
table(df$death2)

```
```{r}
prop.table(table(df$death2))
```

11. Find the first quartile (Q1), second quartile (Q2) and and third quartile (Q3) of the variable `death`. (Hint:  Use the `summary` function)
```{r}
summary(df$Deaths)
```

12. Create the categorical variable `death3` variable taking the values as follows

- `low_death` if the number of deaths smaller than the 25 percentile (Q1)

- `mid_death` if the number of deaths from Q1 to Q3

- `high_death` if the number of deaths greater than Q3

```{r}
library(dplyr)
df$death3 <- case_when(df$Deaths<1171  ~ 'low_death', 
                       df$Deaths<14912 ~ 'mid_death', 
                       TRUE            ~'high_death')
table(df$death3)

```

13. Find the average cases in Rhode Island in 2021
```{r}
df1 <- df[df$Date>='2021-01-01',] 

df2 = df1[df1$State=='Rhode Island',] 

by(df2$Cases, df2$State, mean)
```

14. Find the median cases by weekdays in Rhode Island in 2021 
```{r}
df$weekdays <- weekdays(df$Date)
df1 = df[df$State=='Rhode Island',]
by(data = df$Deaths, INDICES = df$weekdays, FUN = median)
```

15. Compare the median cases in Rhode Island in June, July, August and September in 2021. 
```{r}
df$months <- months(df$Date)
df5 = df[df$State=='Rhode Island',]
df7 = df[(df$Date>="2021-06-01")&(df$Date<="2021-09-30"),]
by(data = df7$Cases, INDICES = df7$months, FUN = median)
```

16.  Find your own dataset, import it and implement the following functions on the data

- head
- str
- names
- mean, min, max
- table
- is.na
- colSums
- class
- cor
- by
- ifelse
- case_when
```{r}
df = read.csv("https://bryantstats.github.io/math421/data/titanic.csv")
head(df)
str(df)
names(df)
summary(df)
table(df$Ticket)
#for the is.na function I used sum(is.na) since the is.na function gave a huge set of data.
sum(is.na(df))
colSums(is.na(df))
class(df)
cor(df$Age, df$Survived)
by(df$Survived, df$Pclass, mean)
df$Survived2 <- ifelse(df$Survived==0, 'survived', 'did_not_survive')
library(dplyr)
df$Survived2 <- case_when(df$Survived==0 ~ 'survived', TRUE ~ 'did_not_survive')
table(df$Survived2)

```

17.  In the dataset in #16, practice the follows.  You can reuse the code of 16. 

- Create a categorical variable from a continuous variable
- From one categorical variable, create a new categorical variable with fewer categories

If you do not have a data, you can use [the titanic dataset](https://www.kaggle.com/competitions/titanic/overview), which can be downloaded at [this link](../data/titanic.csv)
```{r}
df$Age2 <- ifelse(df$Age==0, 'child', 'adult')
library(dplyr)
df$Age2 <- case_when(df$Age==0 ~ 'child', TRUE ~ 'adult')
df$Age2 <- case_when(df$Age<10 ~ 'child', df$Age<25 ~ 'teenager', TRUE~'adult')
table(df$Age2)
```

