
---
output: # rmarkdown::github_document
  html_document:
  pdf_document: default
  word_document: default
title: "Assignment 4 - Data Wrangging with `dplyr`"
---

***How to do it?***: 

- Open the Rmarkdown file of this assignment ([link](assignment4.Rmd)) in Rstudio. 

- Right under each **question**, insert  a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question. 

- Notice that if there is `eval=FALSE` in the first line of the code chunk, the chunk will not be execute.

- `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.  

-  Publish the html file to your Githiub Page. 

***Submission***: Submit the link on Github of the assignment to Canvas under Assignment 4. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```

-------

#### 1. Install `tidyverse` package

An R package can be installed by `install.packages` function. Install `tidyverse` if you have not done so. 

```{r}
library(tidyverse)

```

-------

#### 2. Read the data using `read_csv`

Use `read_csv` function to import the US Covid 19 data at [link](https://covidtracking.com/data/download/all-states-history.csv).  Don't forget to import `tidyverse` (library(tidyverse)) so that you can use read_csv. 
```{r}
library(tidyverse)
df <- read_csv('https://covidtracking.com/data/download/all-states-history.csv')
df

```

-------

#### 3.  Fix the date and ceate some new variables

`lubridate` is a package of the `tidyverse` packages.  We will make uses of `lubridate` in this question.

- Use the below codes to create `month`, `weekday` and `monthday` variables

```{r}
library(lubridate)
df$month = month(df$date)

# day of the week
df$weekday = wday(df$date)

# day of the month
df$monthday <- mday(df$date)
```

-------

#### 4. Create new variables with `case_when`. 

The function `case_when` is a good option to create a new variable from existing variable.  For example, this below codes create a new variable, `daily_death`, from `deathIncrease` variable.  `deathIncrease` is the number of daily new death by Covid19.  The new variable `daily_death` takes three values:  low (if `deathIncrease` less than 3), medium (`deathIncrease` from 3 to 14), and high (`deathIncrease` more than 14).  Please notice that this can also be done in a different way as shown in Assignment 3. 

```{r}
df$daily_death <- case_when(
  df$deathIncrease <3 ~ 'low',
  df$deathIncrease <=14 ~ 'medium',
  TRUE ~ 'high')
```

- Create variable `month2` that takes three values: early_month (day of the month from 1-10), mid_month (day of the month from 11-20), and end_month (day of the month > 20).

- Create variable `weekend` that takes two values:  1 if it's Saturday or Sunday or 0 otherwise. 

```{r}
df$month2 <- case_when(df$monthday < 10 ~ 'early_month', 
                       df$monthday <=20 ~ 'mid_month', 
                       TRUE ~ 'end_month')

```


```{r}
df$weekend <- case_when(df$weekday ==6 ~ 1,
                        df$weekday ==7 ~ 1,
                        TRUE ~ 0)
```

-------

#### 5. Select function

Use the select function to deselect the column `totalTestsViral` from the data.
```{r}
df = df %>% select(-totalTestsViral)
```

-------

#### 6. Pipe Operator ( %>% )

Pipe operator offers another way to write R codes. Many times, it makes the codes more readable.  Pipe works very well with all the `tidyverse` packages. Refer to [these slides](../slides/4_dplyr.html#15) (slide 15, 16, 17 and 18) to rewrite the below codes using pipe operator

```{r}
x <- c(1:10)

# square root of x
sqrt(x)

sum(sqrt(x))

log(sum(sqrt(x)))

# log base 2 of 16
log(16, 2)

```
```{r}
#Rewritten Code with Pipe
x <- c(1:10)

x %>% sqrt

x %>% sqrt %>% sum

x %>% sqrt %>% sum %>% log

16 %>% log(2)
```

-------

#### 7. Combo 1: group_by + summarise

This combo is used when you want to apply a function/calculation to different groups of the data. For example, to calculate the average number of cases (`positiveIncrease`) by `dataQualityGrade`, we use:

```{r}
df %>% 
  group_by(weekday) %>% 
  summarise(mean(positiveIncrease))
```

- Calculate the median number of cases (`positiveIncrease`) by `month`

```{r}
df %>% 
  group_by(month) %>% 
  summarise(median(positiveIncrease))

```
- Calculate the average number of cases (`positiveIncrease`) by `month2`

```{r}
df %>% 
  group_by(month2) %>% 
  summarise(mean(positiveIncrease))
```
- Calculate the median number of cases (`positiveIncrease`) by `weekend`

```{r}
df %>% 
  group_by(weekend) %>% 
  summarise(median(positiveIncrease))
```

-------

#### 8. Combo 2: filter + group_by + summarise

An example: to calculate the average number of cases (`positiveIncrease`) in January and February separately, we use:

```{r}
df %>% 
  filter(month==1|month==2) %>% 
  group_by(month) %>% 
  summarise(positve_increase = mean(positiveIncrease))
```

- Calculate the median number of cases (`positiveIncrease`) on the weekend by `month` in October and November 2020. 

```{r}
df %>% 
  filter(date>='2020-10-01', date<='2020-11-30') %>% 
  group_by(weekend) %>%
  summarise(positive.increase = median(positiveIncrease))
```

- Calculate the average number of death at different periods of a month (`month2` variable) in Fall 2020

```{r}
df %>% 
  filter(date>='2020-09-01', date<='2020-11-30') %>% 
  group_by(month2) %>% 
  summarise(average_death = mean(death))
```

- Compare the average number of hospitalizations between weekdays and weekends in Summer 2020
```{r}
df %>% 
  filter(date>= '2020-06-01', date<= '2020-08-31') %>% 
  group_by(weekend) %>% 
  summarise(Avg_hospitalized = mean(hospitalized, na.rm = TRUE))
```

- Redo Questions 14 and 15 in [Assignment 3](assignment3.html) using the combos.  Notice:  you also need to use the data used in Assignment 3. 


14. Find the median cases by weekdays in Rhode Island in 2021 

```{r}
df2 = read_csv('https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-states.csv')

df2$weekdays <- weekdays(df2$date)

df2 %>% 
  filter(date>='2021-01-01',date<='2021-12-31', state=="Rhode Island") %>%
  group_by(weekdays) %>% 
  summarise(median_cases = median(cases))
```
15. Compare the median cases in Rhode Island in June, July, August and September in 2021. 

```{r}
df2$month <- month(df2$date)

df2 %>% 
  filter(date>='2021-06-01',date<='2021-09-30') %>% 
  group_by(month) %>% 
  summarise(median.cases = median(cases))
```
-----

#### 9. Combo 3: filter + group_by + summarise + arrange

Use the arrange function to find a month that has the highest number of deaths on the weekend.

```{r}
df %>% 
  filter(weekend==1) %>% 
  group_by(month) %>% 
  summarise(max_deaths = sum(death, na.rm = TRUE)) %>% 
  arrange(-max_deaths)
```

----
#### 10. Use your own dataset and implement the follows functions or combos. You can use the [Adult Census Income](../data/adult_census.csv) or [Titanic data](../data/titanic.csv). 

- select
- filter
- mutate
- summarise
- arrange
- count
- count + arrange
- filter + count + arrange
- group_by + summarise
- filter + group_by + summarise
- filter + group_by + summarise + arrange

```{r}
df3 = read_csv("https://bryantstats.github.io/math421/data/adult_census.csv")

df3 %>% select(workclass,education)

```

```{r}
df3 %>% filter(workclass=='Private',age>=30)
```

```{r}
df3 %>% mutate(gain_minus_loss = capital.gain - capital.loss)
```
```{r}
df3 %>% summarise(mean_age = mean(age, na.rm=TRUE))
```
```{r}
df3 %>% arrange(workclass)
```
```{r}
df3 %>% count(workclass)
```

```{r}
df3 %>% 
  count(workclass) %>% 
  arrange(n)
```
```{r}
df3 %>% 
  filter(age>'30') %>%
  count(workclass) %>% 
  arrange(n)
  
  
```
```{r}
df3 %>% 
  group_by(occupation) %>% 
  summarise(mean_gain = mean(capital.gain))
```
```{r}
df3 %>% 
  filter(age>='30') %>% 
  group_by(occupation) %>% 
  summarise(mean_loss = mean(capital.loss))
```

```{r}
df3 %>% 
  filter(capital.gain>='5000') %>% 
  group_by(occupation) %>% 
  summarise(avg.age = mean(age)) %>% 
  arrange(-avg.age)
```
