
---
output: 
  html_document:
  pdf_document: default
  word_document: default
title: "Assignment 9: Visualization - Revealing a graph"
---

***How to do it?***: 

- Open the Rmarkdown file of this assignment ([link](assignment9.Rmd)) in Rstudio. 

- Right under each **question**, insert  a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question. 

- `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.  

-  Publish the html file to your Githiub Page. 

***Submission***: Submit the link on Github of the assignment to Canvas

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```


-------

**Notice:** *In this assignment, all the plot should have title, caption, and axis labels. *

1. Use the dataset by of covid 19 by WHO at https://covid19.who.int/WHO-COVID-19-global-data.csv. Find the three countries with the most numbers of deaths by Covid-19. 
```{r}

library(gganimate)
library(gifski)
library(gapminder)
library(ggplot2)
library(tidyverse)
library(lubridate)
library(knitr)
df <- read_csv("https://covid19.who.int/WHO-COVID-19-global-data.csv")
```

Hint: 

-  Use `filter` to select the most current date 

-  Pipe with `arrange(-Cumulative_deaths)` to rank the data by the number of cumulative deaths

```{r}
df %>% 
  filter(Date_reported == max(Date_reported, na.rm = TRUE),) %>% 
  arrange(-Cumulative_deaths) %>% 
  slice(1:3)
 
```

2. Use `transition_reveal` to make a plot revealing the number of deaths in the three countries with the most numbers of deaths

```{r}
df %>% 
  filter(Country %in% c('United States of America', 'Brazil', 'India')) %>%
  group_by(Country, Date_reported) %>% 
  summarise(totaldeaths = sum(Cumulative_deaths)) %>% 
ggplot(aes(y= totaldeaths, x= Date_reported, color= Country))+
  geom_line()+
  geom_point(size=3)+
  geom_text(aes(label = Country),
            hjust = -.1, size = 5)+
  transition_reveal(Date_reported)
```

3. Create the new variable (column) `death_per_cases` recording the number of deaths per cases (Hint: divide cumulative deaths by cumulative cases).  What are the three countries with the highest deaths per cases? 

```{r}
df = df %>% 
  mutate(death_per_cases = Cumulative_deaths/Cumulative_cases) 

df %>% 
  arrange(-death_per_cases) %>% 
            slice(1:3)

```


4. Use `transition_reveal` to make a plot revealing the number of deaths per cases of the US, Italy and Mexico. 
```{r, warning=FALSE}
df %>% 
  filter(Country %in% c('United States of America', 'Italy', 'Mexico')) %>%
ggplot(aes(y=death_per_cases , x= Date_reported, color= Country))+
  geom_line()+
  geom_point(size=3)+
  geom_text(aes(label = Country),
            hjust = -.1, size = 5)+
  transition_reveal(Date_reported)
```

5. Import the following data https://raw.githubusercontent.com/owid/covid-19-data/master/public/data/vaccinations/vaccinations.csv
               
Use `transition_reveal` to plot the total vaccinations per hundred by level of income. Hint: the levels of income are identified in the `location` variable. Put a short comment on the caption of the plot. 

```{r}
df1 <- read_csv('https://raw.githubusercontent.com/owid/covid-19-data/master/public/data/vaccinations/vaccinations.csv')

df1 %>% 
  filter(location %in% c('High income', 'Low income', 'Lower middle income', 'Upper middle income')) %>% 
  ggplot(aes(y=total_vaccinations, x=date, color= location))+
  geom_line()+
  geom_point(size=3)+
  geom_text(aes(label = location),
            hjust = -.1, size = 5)+
  transition_reveal(date)
  
```

