
---
output: 
  html_document:
  pdf_document: default
  word_document: default
title: "Assignment 8: Visualization - Making a bar race"
---

***How to do it?***: 

- Open the Rmarkdown file of this assignment ([link](assignment8.Rmd)) in Rstudio. 

- Right under each **question**, insert  a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question. 

- `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.  

-  Publish the html file to your Githiub Page. 

***Submission***: Submit the link on Github of the assignment to Canvas

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```


-------

1. Install two packages `gganimate` and `gifski` then restart Rstudio. Using the `Adult Census Income` data, make an animation using `geom_point` and `transition_states`.
```{r}
library(gganimate)
library(gifski)
library(gapminder)
library(ggplot2)
library(tidyverse)
library(lubridate)
library(knitr)

df <- read_csv("https://bryantstats.github.io/math421/data/adult_census.csv")
```

```{r}
g <- ggplot(df,
            aes(x = age,
                y = hours.per.week))+
  geom_point(size=4)+
  transition_states(age)
animate(g)
```


2. Using the `Adult Census Income` data, make an animation using `geom_bar` and `transition_states`.
```{r}
g2 <- ggplot(df, aes(x=race,
                     y=marital.status,
                     fill=marital.status))+
  geom_col()+
  transition_states(age)
animate(g2)
```

3. Use the WHO's dataset at [this link](https://covid19.who.int/WHO-COVID-19-global-data.csv). Make a top-10 bar race by months between countries on the number of deaths by Covid 19 in 2021.
```{r}
library(lubridate)
library(gganimate)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(knitr)

df1 <- read_csv("https://covid19.who.int/WHO-COVID-19-global-data.csv")
df1 %>% filter(Date_reported>= "2021-01-01", Date_reported<="2021-12-31")
df1$month <- month(df1$Date_reported)
d1 <- df1 %>% group_by(month,Country) %>% summarise(mean= mean(Cumulative_deaths))
d2 <- d1 %>% group_by(month) %>% mutate(rank= rank(-mean))
d3 <- d2 %>% filter(rank <= 10)

a1 <- d3 %>% ggplot(aes(x=rank, y=mean, group=Country, fill=Country, label=Country)) + geom_col()+ 
  geom_text(aes(y=mean, label=Country), hjust = 1.4)+
  coord_flip(clip = 'off', expand = FALSE)+scale_x_reverse()+
  labs(title = 'Month {closest_state}', x='', y='Total Number of Deaths', fill='Country')+
  theme(plot.title = element_text(hjust = 1, size = 22), axis.ticks.y = element_blank(),
        axis.text.y = element_blank())+
  transition_states(month)+
  ease_aes('cubic-in-out')

animate(a1, nframes=400)
```

4. Make a bar race using a dataset of your own interest. You may use the dataset that we use in class (https://covidtracking.com/data/download/all-states-history.csv) but you should make a different bar racev from ones in the slides. 
```{r}
df1 <- read_csv("https://covid19.who.int/WHO-COVID-19-global-data.csv")

df1 %>% filter(Date_reported>= "2022-01-01", Date_reported<="2022-12-31")
df1$month <- month(df1$Date_reported)
d1 <- df1 %>% group_by(month,Country) %>% summarise(mean= mean(Cumulative_cases))
  
 d2 <- d1 %>% group_by(month) %>% mutate(rank= rank(-mean))
d3 <- d2 %>% filter(rank<=5)

 a1 <- d3 %>% ggplot(aes(x=rank, y=mean, group=Country, fill=Country, label=Country)) + geom_col()+ 
   geom_text(aes(y=mean, label=Country), hjust = 1.4)+
   coord_flip(clip = "off", expand = FALSE)+scale_x_reverse()+
   labs(title = "Month {closest_state}", x="", y="Total Number of Cases", fill="Country")+
  theme(plot.title = element_text(hjust = 1, size = 22), axis.ticks.y = element_blank(),
        axis.text.y = element_blank())+
  transition_states(month)+
  ease_aes("cubic-in-out")
 animate(a1, nframes=400)
```

